---
title: "Statistical Inference Project_Part 2"
author: "Moh Ahmed"
date: "06/02/2017"
output: html_document
---

# Overview

This report analyses the toothgrowth dataset in the R datasets packages. The aim is to perform basic exploratory data analyses and provide a basic summary of the data. Then, using confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

In this dataset, the response variable is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, either orange juice or ascorbic acid. 

# Summary of the data

ToothGrowth data is a data frame with 60 observations on 3 variables.

[,1]	len	numeric	Tooth length
[,2]	supp	factor	Supplement type (VC or OJ).
[,3]	dose	numeric	Dose in milligrams/day

```{r, warning=FALSE}
library(ggplot2)
library(datasets)
library(gridExtra)

tooth_growth_dt <- ToothGrowth
str(tooth_growth_dt)
summary(tooth_growth_dt)

```


# Exploratory Data Analyses


```{r, warning=FALSE}
tooth_growth_dt <- data.frame(len = tooth_growth_dt$len, supp = tooth_growth_dt$supp, dose = as.factor(tooth_growth_dt$dose))

plot1 <- ggplot(tooth_growth_dt, aes(x=dose, y=len, fill = dose))+ geom_boxplot() + facet_grid(.~tooth_growth_dt$supp) + ggtitle( 'Tooth Length vs Dose vs Supplement') + labs(fill='Dose (mg)') + xlab('Dose (mg)') + ylab('Tooth Length (mm)') + theme(legend.position = 'none')  +  theme(plot.title = element_text( size=12, hjust=0)) + theme(axis.title = element_text( size=8)) 

plot2 <- ggplot(tooth_growth_dt, aes(x=dose, y=len, fill = dose))+ geom_boxplot() +  ggtitle('Tooth Length vs Dose') + labs(fill='Dose (mg)') + xlab('Dose (mg)') + ylab('Tooth Length (mm)') +  theme(plot.title = element_text( size=12, hjust=0)) + theme(axis.title = element_text( size=8))

plot3 <- ggplot(tooth_growth_dt, aes(x=supp, y=len, fill = supp))+ geom_boxplot() + facet_grid(.~tooth_growth_dt$dose) + ggtitle('Tooth Length vs Supplement vs Dose') + labs(fill='Supplement Type') + xlab('Supplement Type') + ylab('Tooth Length (mm)') + theme(legend.position = 'none') +  theme(plot.title = element_text( size=12, hjust=0)) + theme(axis.title = element_text( size=8))

plot4 <- ggplot(tooth_growth_dt, aes(x=supp, y=len, fill = supp))+ geom_boxplot()  + ggtitle('Tooth Length vs Supplement') + labs(fill='Supplement Type') + xlab('Supplement Type') + ylab('Tooth Length (mm)') +  theme(plot.title = element_text( size=12, hjust=0)) + theme(axis.title = element_text( size=8))

grid.arrange(plot1,plot2,plot3,plot4, nrow= 2, ncol = 2)

```


```{r, warning=FALSE}
vc_data <- tooth_growth_dt[tooth_growth_dt$supp=='VC',]
oj_data <- tooth_growth_dt[tooth_growth_dt$supp=='OJ',]

t.test(vc_data$len, oj_data$len)
```




